### 2013-12-03 paper/main.tex
shifting/scaling unnecessary?
It appears that the shift and scaling of the model so
that it accommodates a step function may be unnecessary.
Systems designed from specs have no such requirement.

@@

% {{{ abstract
%\pagebreak
%\thispagestyle{empty}
\begin{abstract}
\noindent
The task of maintaining a stable idle for an internal combustion
engine with spark ignition is non-trivial.
Any time an accessory is turned on/off the torque applied to
the engine changes.
And changes in torque will change the engine rpm if the control
inputs are constant.
This paper shows how control systems methods can be applied to
the problem of idle stabilization.
Methods include: pole placement, direct design, and
various state space designs.

@@

There is yet another complication with this model.
Typical systems are in a normalized so that it is possible to
achieve an output of one given a step input.
With this engine model an output of 1 rpm is an impossibly slow value.
And an input of zero or one rpm are similarly impossible.

To resolve this issue the values will be scaled in to a normal range.
A normal idle rpm range is from 600 to 1200 rpm.
To scale these values so that zero equals 600 and one equals 1000
Equation \ref{eq:scale} can be used.

\begin{align}
	y &= \frac{1}{400}x - \frac{3}{2} \label{eq:scale}
\end{align}

Figure \ref{fig:stable} shows the stable rpms for a given control input.
Its behavior is described by Equation \ref{eq:stable}.

\begin{align}
	y &= 1450.1 x \label{eq:stable}
\end{align}

% {{{ fig:stable
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.6]{../octave/stable_rpm_plot}
\end{center}
\caption{Stable rpm reached for a given constant control input.}
\label{fig:stable}
\end{figure}
% }}}

With these normalization corrections applied the response can be
seen in Figure \ref{fig:normeng}.
To avoid the zero rpm case it is allowed to stabilize before the
step is applied.
The lowest plot shows the response after the step.
It takes slightly over one second to stabilize, which is better than
the inaccurate case from zero rpm, but it is still slow and can
be improved.

% {{{ fig:normeng
\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{../octave/norm_engine_model_plot}
\includegraphics[scale=0.55]{../octave/norm_engine_plot_close}
\end{center}
\caption{Normalized engine model response.  After it stabilizes
at unit step is applied to produce the normalized response.
The lowest plot is a close up of the step response in the center
plot.}
\label{fig:normeng}
\end{figure}
% }}}

>>>>

% prevent breaks in the middle of paragraphs
\widowpenalties 1 10000
\raggedbottom

@@

% {{{ abstract
%\pagebreak
%\thispagestyle{empty}
\begin{abstract}
\noindent
The task of maintaining a stable idle for an internal combustion
engine with spark ignition is non-trivial.
Any time an accessory is turned on/off the torque applied to
the engine changes.
And changes in torque will change the engine rpm if the control
inputs are constant.
This paper shows how control system methods can be applied to
the problem of idle stabilization.
Because the engine model is inherently discrete all of the methods
used are also discrete.
Methods include: pole placement, direct design, and
various state space designs.

@@
